---
title: "knb-lter-cap.632"
author: "CAP LTER"
---

# README

content moved to project README.md

# annuals biomass

```{r annuals-biomass, eval=TRUE}

annuals_biomass <- dbGetQuery(pg,"
SELECT
  s.code as site_code,
  ab.plot_id,
  t.code as treatment_code,
  ab.location_within_plot as location_within_plot,
  ab.replicate as subplot,
  ab.subquad_orientation,
  ab.date,
  ab.year,
  ab.mass,
  ab.notes
FROM
  urbancndep.annuals_biomass ab
  JOIN urbancndep.plots p ON (ab.plot_id = p.id)
  JOIN urbancndep.sites s ON (s.id = p.site_id)
  JOIN urbancndep.treatments t ON (t.id = p.treatment_id)
ORDER BY year, plot_id, location_within_plot, subplot, subquad_orientation;")

# change column types as appropriate
annuals_biomass <- annuals_biomass |>
  dplyr::mutate(
    site_code            = as.factor(site_code),
    treatment_code       = as.factor(treatment_code),
    location_within_plot = as.factor(location_within_plot),
    subplot              = as.factor(subplot),
    plot_id              = as.character(plot_id)
  )

# write_attributes(annuals_biomass)
# write_factors(annuals_biomass)

annuals_biomass_desc <- "Biomass (g) of annual plants harvested from subplots within Desert Fertilization study plots. One-meter subplots include locations around a Larrea tridentata plant and locations in the interplant space between shrubs. Material is harvested from 0.25 square meter quadrats within each subplot. All harvests occur during the spring."

# create data table based on metadata provided in the companion csv
annuals_biomass_DT <- capeml::create_dataTable(
  dfname         = annuals_biomass,
  description    = annuals_biomass_desc,
  dateRangeField = "date",
  overwrite      = TRUE
)

```

# annuals composition

```{r annuals-composition}

# add filter to remove 2008 data from publication
# add filter to remove total comparable annual cover from publication (could be removed from DB as well)

annuals_composition <- dbGetQuery(pg,"
SELECT
  s.code as site_code,
  ce.plot as plot_id,
  t.code as treatment_code,
  ce.patch_type as location_within_plot,
  ce.subplot,
  ce.sample_date as date,
  ce.year,
  ce.collector,
  ct.cover_type,
  ct.cover_category,
  cc.cover_amt as cover_amount
FROM
  urbancndep.cover_composition cc
  JOIN urbancndep.cover_events ce ON (cc.cover_event_id = ce.cover_event_id)
  JOIN urbancndep.cover_types ct ON (cc.cover_type_id = ct.cover_type_id)
  JOIN urbancndep.plots p ON (ce.plot = p.id)
  JOIN urbancndep.sites s ON (s.id = p.site_id)
  JOIN urbancndep.treatments t ON (t.id = p.treatment_id)
WHERE
  ce.year > 2008 AND
  ct.cover_type != 'total_comparable_annual_cover'
ORDER BY year, plot, location_within_plot, subplot, cover_type;
")

# change column types as appropriate
annuals_composition <- annuals_composition |>
  dplyr::mutate(
    site_code            = as.factor(site_code),
    plot_id              = as.character(plot_id),
    treatment_code       = as.factor(treatment_code),
    location_within_plot = as.factor(location_within_plot),
    subplot              = as.factor(subplot),
    cover_type           = as.factor(cover_type),
    cover_category       = as.factor(cover_category)
  )

# write_attributes(annuals_composition)
# write_factors(annuals_composition)

annuals_composition_desc <- "Composition of annual plants and some other characteristics (e.g., bare soil, base or canopy of perennial plants) at subplots within Desert Fertilization study plots. One-meter subplots include locations around a Larrea tridentata plant and locations in the interplant space between shrubs. All measurements collected in the spring."

# create data table based on metadata provided in the companion csv
# do not add date range since early years did not have a full date
annuals_composition_DT <- capeml::create_dataTable(
  dfname      = annuals_composition,
  description = annuals_composition_desc,
  overwrite   = TRUE
)

```

# fertilizer

```{r fertilizer, eval=TRUE}

fertilizer_application <- dbGetQuery(pg,'
SELECT
  s.code as site_code,
  f.date as application_date,
  f."N" as nitrogen,
  f."P" as phosphorus
FROM urbancndep.fertilizer_applications f
  JOIN urbancndep.sites s ON (f.site_id = s.id)
ORDER BY f.date, s.code
;')

# change column types as appropriate
fertilizer_application <- fertilizer_application |>
  dplyr::mutate(site_code = as.factor(site_code))

# write_attributes(fertilizer_application)
# write_factors(fertilizer_application)

fertilizer_application_desc <- "catalog of amounts and timing of nitrogen and phosphorus fertilizer applications to nitrogen (N), phosphorus (P), and nitrogen+phosphorus (N+P) treatment plots - applications are delivered to the respective plot at the site that receives either N, P, or an N+P addition, except control sites (n = 2) that do not receive a fertilizer amendment"

fertilizer_application_DT <- capeml::create_dataTable(
  dfname         = fertilizer_application,
  description    = fertilizer_application_desc,
  dateRangeField = "application_date",
  overwrite      = TRUE
)

```

# stems

```{r stems, eval=TRUE}

stem_growth <- dbGetQuery(pg,"
SELECT
  s.code AS site_code,
  p.id AS plot_id,
  t.code AS treatment_code,
  sp.scientific_name,
  sh.code AS shrub_code,
  st.direction,
  st.pre_date,
  st.post_date,
  st.post_note,
  sl.post_measurement,
  sl.length_in_mm AS stem_length,
  sc.comment AS stem_comment,
  agg_plot_notes.plot_comment
FROM urbancndep.stems st
JOIN urbancndep.shrubs sh ON sh.id = st.shrub_id
JOIN urbancndep.plots p ON sh.plot_id = p.id
JOIN urbancndep.sites s ON p.site_id = s.id
JOIN urbancndep.treatments t ON p.treatment_id = t.id
JOIN urbancndep.shrub_species sp ON sh.shrub_species_id = sp.id
LEFT JOIN urbancndep.stem_lengths sl ON st.id = sl.stem_id
LEFT JOIN urbancndep.stem_comment sc ON (sc.stem_id = st.id AND sl.post_measurement = sc.post_measurement)
LEFT JOIN (
  SELECT
    plot_id,
    survey_date,
    string_agg(distinct plot_notes, '; ') AS plot_comment
  FROM urbancndep.stem_plot_notes
  GROUP BY plot_id, survey_date
) agg_plot_notes ON (agg_plot_notes.plot_id = p.id AND agg_plot_notes.survey_date = st.post_date)
WHERE
NOT (
    -- omit post data not collected when smart sampling was implemented
    EXTRACT (YEAR FROM st.pre_date) = 2016 AND
    EXTRACT (MONTH FROM st.pre_date) = 10 AND
    st.post_date IS NULL
  ) AND
  NOT (p.id = 13 AND st.pre_date = '2010-05-10') AND
  NOT (p.id = 12 AND st.pre_date = '2010-05-11')
ORDER BY
  st.pre_date,
  p.id,
  sl.post_measurement,
  sh.code,
  st.direction;
")

stem_growth <- stem_growth |>
  dplyr::mutate(
    site_code        = as.factor(site_code),
    treatment_code   = as.factor(treatment_code),
    shrub_code       = as.factor(shrub_code),
    direction        = as.factor(direction),
    plot_id          = as.character(plot_id),
    post_measurement = as.factor(post_measurement),
    post_note        = gsub("[\r\n]", "", post_note)
  )

# write_attributes(stem_growth)
# write_factors(stem_growth)

stem_growth_desc <- "Biannual measures of stem growth on five Larrea tridentata study plants in Desert Fertilization experiment treatment and control plots"

stem_growth_DT <- capeml::create_dataTable(
  dfname         = stem_growth,
  description    = stem_growth_desc,
  dateRangeField = "pre_date",
  overwrite      = TRUE
)

```

# PRS

```{r PRS, eval=TRUE}

plant_root_simulator <- dbGetQuery(pg,'
SELECT
   s.code AS site_code,
   p.id AS plot_id,
   t.code AS treatment_code,
   pa.start_date,
   pa.end_date,
   pa.analyte,
   pa.final_value,
   pa.flag,
   pa.location_within_plot,
   pa.num_cation_probes,
   pa.num_anion_probes
 FROM urbancndep.sites s
   JOIN urbancndep.plots p ON s.id = p.site_id
   JOIN urbancndep.treatments t ON p.treatment_id = t.id
   JOIN urbancndep.prs_analysis pa ON p.id = pa.plot_id
ORDER BY pa.start_date;')

# standardize sample type names
plant_root_simulator <- plant_root_simulator |>
  dplyr::filter(!grepl("NH4", location_within_plot)) |> # remove the washing experiment samples
  dplyr::mutate(
    location_within_plot = replace(location_within_plot, location_within_plot == "between plants", "between_plant"),
    location_within_plot = tolower(location_within_plot),
    location_within_plot = gsub(" ", "_", location_within_plot),
    location_within_plot = gsub("-", "_", location_within_plot),
    analyte              = gsub(" ", "-", analyte)
  )

# change column types as appropriate
plant_root_simulator <- plant_root_simulator |>
  dplyr::mutate(
    site_code            = as.factor(site_code),
    treatment_code       = as.factor(treatment_code),
    analyte              = as.factor(analyte),
    location_within_plot = as.factor(location_within_plot),
    plot_id              = as.character(plot_id)
  )

# write_attributes(plant_root_simulator)
# write_factors(plant_root_simulator)

plant_root_simulator_desc <- "Soil ion concentrations as determined with Plant Root Simulator (PRS®) probes (ion exchange resin membranes). Probes for the analyses of soil anions have a positively-charged membrane to simultaneously attract and adsorb all negatively-charged anions, such as nitrate (NO3-), phosphate (H2PO4-, HPO42-), and sulphate (SO42-), whereas cation probes have a negatively-charged membrane to simultaneously attract and adsorb all positively-charged cations, such as ammonium (NH4+), potassium (K+), calcium (Ca2+), and magnesium (Mg2+)."

plant_root_simulator_DT <- capeml::create_dataTable(
  dfname         = plant_root_simulator,
  description    = plant_root_simulator_desc,
  dateRangeField = "start_date",
  overwrite      = TRUE
)

```

# study locations

Here we keep the source file (desert_fertilization_sampling_sites.kml) in the
repository so that we can easily reconstruct desert_fertilization_sites when
updating.

```{r generalized-study-locations, eval=TRUE}

desert_fertilization_sampling_sites <- sf::read_sf(
  dsn = "desert_fertilization_sampling_sites.kml"
  ) |>
dplyr::mutate(description = dplyr::case_when(
    Name == "MCS" ~ "McDowell Mountain Park south",
    Name == "SMW" ~ "South Mountain Park west",
    Name == "WTM" ~ "White Tanks Park",
    Name == "DBG" ~ "Desert Botanical Garden",
    Name == "LDP" ~ "Lost Dutchman State Park",
    Name == "MVP" ~ "Mountain View Park",
    Name == "PWP" ~ "Piestewa Peak",
    Name == "SRR" ~ "Salt River Recreation Area",
    Name == "MCN" ~ "McDowell Mountain Park north",
    Name == "UMP" ~ "Usury Mountain Park",
    Name == "SNE" ~ "Sonoran National Monument east",
    Name == "EMW" ~ "Estrella Mountain Park west",
    Name == "EME" ~ "Estrella Mountain Park east",
    Name == "SME" ~ "South Mountain Park east",
    Name == "SNW" ~ "Sonoran National Monument west"
  )
) |>
purrr::discard(~all(is.na(.x))) |>
dplyr::select(
  site = Name,
  description,
  geometry
)

try(capeml::write_attributes(desert_fertilization_sampling_sites, overwrite = FALSE))

desert_fertilization_sampling_sites_desc <- "approximate location of desert fertilization long-term study sites"

desert_fertilization_sampling_sites_SV <- capemlGIS::create_vector(
  vector_name   = desert_fertilization_sampling_sites,
  description   = desert_fertilization_sampling_sites_desc,
  driver        = "GeoJSON",
  projectNaming = TRUE,
  overwrite     = TRUE
)

```

# tissue CHN

```{r tissue CHN}

chnDataQuery <- '
  SELECT
    site_code,
    plot_id, 
    treatment_code,
    sample_date,
    season_year, 
    tissue_type,
    "Weight",
    "Comment", 
    "Carbon %",
    "Hydrogen %",
    "Nitrogen %"
  FROM urbancndep.plant_tissue_chn
  WHERE 
    plot_id IS NOT NULL AND
    (
    	"Comment" !~* \'need|require\' OR
    	"Comment" IS NULL
    );'

tissue_chn <- dbGetQuery(pg, chnDataQuery)

tissue_chn <- tissue_chn |>
  tidyr::gather(key = analyte, value = percent_composition, `Carbon %`, `Hydrogen %`, `Nitrogen %`) |>
  dplyr::select(
    site_code, plot_id, treatment_code, sample_date, season_year, tissue_type, analyte, Weight, percent_composition, Comment
    ) |>
  dplyr::mutate(
    plot_id        = as.character(plot_id),
    site_code      = as.factor(site_code),
    treatment_code = as.factor(treatment_code),
    tissue_type    = as.factor(tissue_type),
    analyte        = as.factor(analyte)
  ) 

  # write_attributes(tissue_chn)
  # write_factors(tissue_chn) 

tissue_chn_desc <- "CHN (Carbon, Hydrogen, and Nitrogen) elemental analysis of Larrea tridentata leaf tissue and Pectocarya recurvata (whole plant) tissue collected from control plots at Desert Fertilization study sites."

tissue_chn_DT <- capeml::create_dataTable(
  dfname         = tissue_chn,
  description    = tissue_chn_desc,
  dateRangeField = "sample_date",
  overwrite      = TRUE
)

```

# tissue_icp

Owing to the complexity of the workflow and the infrequency of new data of this
type, the initial ICP tissue data processing is isolated in a separate workflow
(`icp_tissue.R`). For DesFert updates that do not include new tissue ICP data,
xml for the tissue icp dt and other entity are recycled from earlier versions
of the knb-lter-cap.632 xml/eml.

```{r tissue-icp, eval=TRUE}

tissue_icp <- read.csv("working_icp/632_tissue_icp_7d940d0979cc34c094c46525f7cd2995.csv") |>
dplyr::mutate(
  season_year = dplyr::case_when(
    season_year == "fall_2010" & sample_date <= "2010-09-01" ~ "spring_2010",
    TRUE ~ season_year 
    ),
  instrument = dplyr::case_when(
    season_year == "fall_2015" & isotope_element == "S_182.0" ~ "ICP-OES",
    season_year == "fall_2015" & isotope_element != "S_182.0" ~ "ICP-MS",
    TRUE ~ instrument
    ),
  isotope_element = dplyr::case_when(
    isotope_element == "S_182.0" ~ "S",
    TRUE ~ isotope_element
    ),
  site_code = dplyr::case_when(
    site_code == "DGB" ~ "DBG",
    site_code == "WHT" ~ "WTM",
    TRUE ~ site_code
  )
  ) |>
dplyr::mutate(
  site_code      = as.factor(site_code),
  plot_id        = as.character(plot_id),
  treatment_code = as.factor(treatment_code),
  sample_date    = as.Date(sample_date),
  tissue_type    = as.factor(tissue_type),
  instrument     = as.factor(instrument)
)

# write_attributes(tissue_icp)
# write_factors(tissue_icp)

tissue_icp_desc <- "Elemental composition of Larrea tridentata leaf tissue and Pectocarya recurvata (whole plant) tissue collected from control plots at Desert Fertilization study sites. Most analyses are by ICP-MS except Sulfur (S), which is typically analyzed by ICP-OES with the instrument type noted in the instrument field."

tissue_icp_DT <- capeml::create_dataTable(
  dfname         = tissue_icp,
  description    = tissue_icp_desc,
  dateRangeField = "sample_date",
  overwrite      = TRUE
)

```

# icp_raw_data

```{r icp-raw-data, eval=TRUE}

icp_raw_data_desc <- "This zipped file contains the raw ICP-MS and ICP-OES data (as XSLM and xls files) pertaining to the analyses of Larrea tridentata leaf tissue and Pectocarya recurvata plant tissue samples. The calculated concentrations are presented in an analysis-friendly format in the data entity 'tissue_icp' that is part of this dataset; the raw data file from which the calculated concentrations were derived is referenced in the source_file field of the tissue_icp data entity."

icp_raw_data_OE <- capeml::create_otherEntity(
  target_file_or_directory = "raw_icp/",
  description              = icp_raw_data_desc,
  overwrite                = TRUE
)

```

# atmospheric deposition

There is a sharp contrast in the structure of the Lachat data where dilutions
are not calculuted in an earlier method but are calculated by the method in
later years. Among the resin data, the transition occurs between 2012 and 2014.
The workflow below included steps to calculate the concentration factoring in
the dilution where needed for earlier data.

Careful that the transition between the Lachat format occurs in 2013 for the
set of data available for knb-lter-cap.632.9 but that may not always be the
case as new data are added from historic runs. For example, it is unclear when
exactly the method changed. This will not be an issue once we have addressed
all of the historic data, but the data should be examined until that time.

We add a unique run_id for each Lachat run. This is needed so that a user can
associate a blanks with the corresponding samples in an analytical run,
espcially since there are not collection dates on blanks. Here I use a simple
integer for the run_id instead of somthing like the Lachat run_file_name from
which the run_id is generated just as something a bit simpler and more
digestible to a user.

For atmospheric deposition blanks:

* solution blanks (e.g., `BLK.BLK1`) are the 0.2 M KCl extract solution that is
  run through a filter. This solution is used to construct the standard curves
  so, really, this information is probably not all that useful to a user since
  you would not want to back out a blank signal from a solution used also for
  the stands but the data are included for completeness.
* field blanks (e.g., `WTM.LATR CNTL1`) are the resin collectors that are
  capped and deployed to the field but never exposed to the atmosphere.
* run blanks (e.g., `resin.KCl.1`) are constructed by performing the complete
  extraction procedure on an aliquot of clean resin material that was not
  deployed in the field or otherwise used in any way.

```{r atmospheric-deposition, eval=TRUE}

allresin <- dbGetQuery(pg, " select * from urbancndep.resin ; ")

unknowns <- allresin |>
dplyr::filter(
  grepl("unknown", sample_type, ignore.case = TRUE),
  omit == FALSE
  ) |>
dplyr::select(
  upload_batch,
  field_id,
  collection_date,
  notes,
  detection_date,
  detection_time,
  manual_dilution_factor,
  auto_dilution_factor,
  analyte_name,
  peak_concentration,
  conc_x_adf,
  conc_x_mdf,
  conc_x_adf_x_mdf,
  sourcefile,
  run_file_name
)

atmospheric_deposition <- dplyr::bind_rows(
  unknowns |>
    dplyr::filter(detection_date < "2013-01-01") |>
    dplyr::rowwise() |>
    dplyr::mutate(max_factor = max(manual_dilution_factor, auto_dilution_factor, na.rm = TRUE)) |>
    dplyr::ungroup() |>
    dplyr::mutate(conc_x_adf_x_mdf = peak_concentration * max_factor) |>
    dplyr::select(-max_factor),
  unknowns |>
    dplyr::filter(detection_date > "2013-01-01")
  ) |>
dplyr::select(
  field_id,
  collection_date,
  notes,
  analyte_name,
  concentration = conc_x_adf_x_mdf,
  run_file_name
  ) |>
# add a unqiue ID for each to accommodate associating samples and blanks
dplyr::left_join(
  unknowns |>
    dplyr::distinct(run_file_name) |>
    tibble::tibble() |>
    dplyr::mutate(run_id = seq_along(run_file_name)),
  by = c("run_file_name")
  ) |>
dplyr::select(
  run_id,
  everything(),
  -run_file_name
  ) |>
dplyr::arrange(
  collection_date,
  field_id,
  analyte_name
  ) |>
dplyr::mutate(
  run_id = as.character(run_id),
  field_id = dplyr::case_when(
    grepl("blk", field_id, ignore.case = TRUE) ~ "KCl_blank",
    TRUE ~ field_id 
    ),
  field_id = as.factor(field_id),
  analyte_name = as.factor(analyte_name)
) |>
dplyr::mutate(
  site_code = stringr::str_extract(field_id, "[A-z]+"),
  site_code = dplyr::case_when(
    grepl("blank", field_id, ignore.case = TRUE) ~ NA_character_,
    grepl("resin", field_id, ignore.case = TRUE) ~ NA_character_,
    TRUE ~ site_code
    ),
  site_code = as.factor(site_code) 
)

try({
  capeml::write_attributes(atmospheric_deposition, overwrite = FALSE)
  capeml::write_factors(atmospheric_deposition, overwrite = FALSE)
})

atmospheric_deposition_desc <- "ammonium-nitrogen and nitrate-nitrogen as measured by ion exchange resin (IER) collectors that are used to measure bulk (wet) deposition in interplant open spaces and throughfall (wet and dry) deposition under the dominant shrub"

atmospheric_deposition_DT <- capeml::create_dataTable(
  dfname         = atmospheric_deposition,
  description    = atmospheric_deposition_desc,
  dateRangeField = "collection_date",
  overwrite      = TRUE
)

```

# people

```{r people, eval=TRUE}

jonAllen <- gioseml::create_role(
  firstName = "jon",
  lastName  = "allen",
  roleType  = "creator"
)

nancyGrimm <- gioseml::create_role(
  firstName = "n",
  lastName  = "grimm",
  roleType  = "creator"
)

sharonHall <- gioseml::create_role(
  firstName = "shar",
  lastName  = "hall",
  roleType  = "creator"
)

jasonKaye <- gioseml::create_role(
  firstName = "jaso",
  lastName  = "kaye",
  roleType  = "creator"
)

creators <- list(
  nancyGrimm,
  sharonHall,
  jasonKaye,
  jonAllen
)

stevanEarl <- gioseml::create_role(
  firstName = "s",
  lastName  = "earl",
  roleType  = "metadata"
)

marisaMasles <- gioseml::create_role(
  firstName = "m",
  lastName  = "masles",
  roleType  = "metadata"
)

quincyStewart <- gioseml::create_role(
  firstName = "q",
  lastName  = "stewart",
  roleType  = "metadata"
)

sallyWittlinger <- gioseml::create_role(
  firstName = "s",
  lastName  = "wittlinger",
  roleType  = "metadata"
)

cook <- gioseml::create_role(
  giosPersonId = 7045,
  #   firstName = "elizabeth",
  #   lastName = "cook",
  roleType     = "metadata"
)

metadataProvider <- list(
  stevanEarl,
  marisaMasles,
  quincyStewart,
  sallyWittlinger,
  cook
)

```

# coverages

```{r coverages}

getMaxDate <- function() {
  
  maxDates <- data.frame()
  maxDates <- rbind(
    dbGetQuery(pg, 'SELECT MAX(sample_date) FROM urbancndep.cover_events;'),
    dbGetQuery(pg, 'SELECT MAX(date) FROM urbancndep.annuals_biomass;'),
    dbGetQuery(pg, 'SELECT MAX(date) FROM urbancndep.fertilizer_applications;'),
    dbGetQuery(pg, 'SELECT MAX(post_date) FROM urbancndep.stems;'),
    dbGetQuery(pg, 'SELECT MAX(end_date) FROM urbancndep.prs_analysis;')
  )
 
  theMaxDate <- max(maxDates$max)
  
  return(theMaxDate)
  
}


geographic_coverage <- split(
  x = desert_fertilization_sampling_sites,
  f = desert_fertilization_sampling_sites$site
) |>
{\(site) purrr::map(.x = site, ~ capeml::create_geographic_coverage(.x,
description = .x$site))}() |> unname()

coverage <- list(
  temporalCoverage = list(
    rangeOfDates = list(
      beginDate = list(
        calendarDate = "2005-12-07"
        ),
      endDate = list(
        calendarDate = as.character(getMaxDate())
      )
    )
  ), # close temporalCoverage
geographicCoverage = geographic_coverage
) # close coverage

# begindate <- "2005-12-07"
# enddate   <- as.character(getMaxDate())
# geographicDescription <- "desert and desert-remnant regional parks in the CAP LTER study encompassing the greater Phoenix metropolitan area and surrounding Sonoran desert region"
# coverage <- EML::set_coverage(
#   begin                 = begindate,
#   end                   = enddate,
#   geographicDescription = geographicDescription,
#   west                  = -112.547375174279352,
#   east                  = -111.482761068522152,
#   north                 = +33.726771147871851,
#   south                 = +33.013262396669028
# )

```

## taxonomic coverage

*Note* that the `taxa_map.csv` built with the `create_taxa_map()` function and
resolving taxonomic IDs (i.e., `resolve_comm_taxa()`) only needs to be run once
per version/session -- the taxonomicCoverage can be built as many times as
needed with `resolve_comm_taxa()` once the `taxa_map.csv` has been generated and
the taxonomic IDs resolved.

```{r taxonomyCleanr, eval=TRUE}

my_path <- getwd() # taxonomyCleanr requires a path (to build the taxa_map)

# get dataset taxa
dataset_taxa <- dbGetQuery(pg,"
  SELECT
  DISTINCT(cover_type)
  FROM urbancndep.cover_types
  WHERE cover_category LIKE 'annual';"
  ) |> 
dplyr::mutate(cover_type = gsub("_", " ", cover_type)) |>
dplyr::filter(!stringr::str_detect(cover_type, "unident")) |> 
# annuals taxa above, additional taxa below
tibble::add_row(cover_type = 'Larrea tridentata') |> 
tibble::add_row(cover_type = 'Ambrosia dumosa') |> 
tibble::add_row(cover_type = 'Ambrosia deltoidea') |> 
tibble::add_row(cover_type = 'Pectocarya recurvata') |> 
tibble::add_row(cover_type = 'Amsinckia menziesii') |> 
tibble::add_row(cover_type = 'Schismus arabicus') |> 
dplyr::distinct(cover_type)

# create or update map. A taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv and put it in the path identified with
# my_path.
taxonomyCleanr::create_taxa_map(path = my_path, x = dataset_taxa, col = "cover_type") 

# resolve taxa by attempting to match the taxon's In this case, data.source 3 is
# ITIS (which, by the way, is the only authority taxonomyCleanr will allow for
# common names).
# resolve_comm_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS
taxonomyCleanr::resolve_sci_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

# build the EML taxonomomic coverage
taxaCoverage <- taxonomyCleanr::make_taxonomicCoverage(path = my_path)

# add taxonomic to other coverages
coverage$taxonomicCoverage <- taxaCoverage

```

# literature cited

```{r literature-cited, eval=TRUE}

# cook <- create_citation("https://doi.org/10.1016/j.envpol.2018.04.013")
fenn   <- capeml::create_citation("https://doi.org/10.2134/jeq2004.2007")
simkin <- capeml::create_citation("https://doi.org/10.1023/B:WATE.0000019958.59277.ed")

citations <- list(
  citation = list(
    fenn,
    simkin 
  ) # close list of citations
) # close citation

# dataset$literatureCited <- citations

```

# dataset

Optionally, provide: scope, abstract, methods, keywords, publication date.
Projects scopes include lter (default), urex, ltreb, and som.

```{r construct-dataset}

dataset <- capeml::create_dataset()
```

# custom units

```{r custom-units, eval=TRUE}

# standardUnits <- get_unitList()
# unique(standardUnits$unitTypes$id) # unique unit types

# the PRS data does include a custom unit but creating this is not required if
# you are recycling previously constructed EML/XML with just updated data, but
# this will be required if you recreate 632 from scratch.

custom_units <- rbind(
  data.frame(
    id             = "microgramPerTenSquareCentimeterPerBurialLength",
    unitType       = "unknown",
    parentSI       = "unknown",
    multiplierToSI = NA,
    description    = "net rate of nutrient ion adsorption by the PRS® Probe expressed as the weight of nutrient adsorbed per surface area of ion-exchange membrane over time"
    ),
  data.frame(
    id             = "milligramPerKilogram",
    unitType       = "massPerMass",
    parentSI       = "gramsPerGram",
    multiplierToSI = 0.000001,
    description    = "millgram of isotope per kilogram of plant tissue")
)

unitList <- EML::set_unitList(
  custom_units,
  as_metadata = TRUE
)

```

# eml

## eml create

```{r construct-eml, eval=TRUE}

eml <- capeml::create_eml()
```

## eml validate

```{r validate-eml, eval=TRUE}

EML::eml_validate(eml)
```

## eml write

```{r eml-to-file, eval=TRUE}

# write the eml to file
capeml::write_cap_eml()
```

# file placement

```{r package-details, eval=TRUE}

# retrieve package details from config.yaml
if (!file.exists("config.yaml")) {
  stop("config.yaml not found")
}
packageIdent <- yaml::yaml.load_file("config.yaml")$packageIdent
packageNum <- yaml::yaml.load_file("config.yaml")$packageNum
```

```{r preview_data_file_to_upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(packageNum, "_"))
```

Move data and final xml files to respective ASU locations.

```{r S3_helper_functions}

# functions and setting for uploading to S3
library(aws.s3)
source("~/Documents/localSettings/aws.s3")
```

```{r upload_data_S3}

# upload files to S3
lapply(list.files(pattern = paste0(packageNum, "_")), gioseml::data_to_amz)
```


# EDI

## EDI: login

```{r edi-login, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::login(
  userId   = keyring::key_get("edi_user"),
  userPass = keyring::key_get("edi_pass")
)

```

## EDI: evaluate

```{r edi-evaluate, eval=TRUE, echo=TRUE, message=TRUE}

evaluation <- EDIutils::evaluate_data_package(
  eml         = paste0(packageIdent, ".xml"),
  useChecksum = TRUE,
  env         = "staging"
)

Sys.sleep(8)

eval_status <- EDIutils::check_status_evaluate(
  transaction = evaluation,
  env         = "staging"
)

if (eval_status) {

  # evaluation summary

  EDIutils::read_evaluate_report_summary(
    transaction = evaluation,
    env         = "staging"
  )

}

# evaluation detailed

# EDIutils::read_evaluate_report(
#   transaction = evaluation,
#   env         = "staging"
# )

```

## EDI: update

```{r edi-update, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::update_data_package(
  eml         = paste0(packageIdent, ".xml"),
  useChecksum = TRUE,
  env         = "production"
)

```

## EDI: logout

```{r edi-logout, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::logout()

```


# post processing

remove data files (if desired)

```{r delete-data-files}

dataFilesToRemove <- dir(pattern = paste0(packageNum, "_"))
file.remove(dataFilesToRemove)
```

XML/EML file to Amazon and cap-metadata

```{r delete-data-files}

gioseml::eml_to_amz(list.files(pattern = "knb.+xml"))

file.copy(list.files(pattern = "knb.+xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")
file.remove(list.files(pattern = "knb.+xml"))
```
